---
title: "btmobile25a power simulation"
author: "Anonymized for peer review (JT)"
reviewer: "Anonymized for peer review (NC)"
output: word_document
editor_options: 
  chunk_output_type: console
---
# Setup
```{r }
# specify data directory
data_dir <- Sys.getenv("data_dir")

# source setup file
source('setup.R')
```

# Open combined data
```{r}
DF <- 
  readRDS(
    file.path(data_dir,
              'combined',
              'DF.full.combined.processed.Rds')
    ) %>% 
  
  # focus on first three years
  filter(response_time < (365*3)) %>% 
  
  # input 0 into the entries that are missing Altmetric data
  mutate(across(c(policy.per, 
                  news.per), 
                ~ replace(., 
                          is.na(.), 
                          0)
                )
         )
```

# Simulation function
Using pilot data, randomly samples N papers and test which model (quadratic or logarithmic) best fits the data.

Returns best fitting model, value, and difference for weighted RMSE and BIC. 
```{r}
# Create simulation function that:
# A: randomly samples the number of authors from the original dataframe 
# B: generates speed, citation, policy and news values based on model parameters

sim <- function(n = 100000, 
                DF = DF){
  
  #######################
  # Step 1: Sample data
  #######################
  sampled_data <- DF %>%
    select(dataset, authors.n, citation.per, 
           news.per, policy.per, speed.per) %>%
    
    # collect sample
    slice_sample(n = n) %>%
    
    # assign weights for later regression
    group_by(authors.n) %>%
    mutate(weight = 1/n()) %>%
    ungroup() 
  
    # Create dataframe that will be populated with best fitting model info
    outcome_df <- data.frame(
      sample_size = numeric(),  # number of sampled papers
      
      speed.quad = numeric(),  # speed quadratic model's p-value
      citation.log = numeric(),  # citation log model's p-value
      policy.quad = numeric(),  # policy quadratic model's p-value
      news.quad = numeric(),  # news quadratic model's p-value
      
      speed.quad.BIC = numeric(),  # speed quadratic model's BIC difference
      citation.log.BIC = numeric(),  # citation log model's BIC difference
      policy.quad.BIC = numeric(),  # policy quadratic model's BIC difference
      news.quad.BIC = numeric())  # news quadratic model's BIC difference

    #######################
    # Step 2: Populate sample size
    #######################
    outcome_df[1,] <- c(n, rep(NA, times = 8))
    
    #######################
    # Step 3: Fit and evaluate speed models
    #######################
    ## linear modeling
    m.s.lin <- lmer(speed.per ~ authors.n + (1|dataset),
                    data = sampled_data,
                    weights = weight)
    
    ### extract weighted residuals
    m.s.lin.res <- 
      #### extract residuals and weights
      data.frame(res = residuals(m.s.lin),
                 weight = m.s.lin@frame$`(weights)`) %>%
      
      #### calculate weighted residuals
      mutate(w.res = sqrt(weight*res^2))
    
    
    ## quadratic modeling
    m.s.quad <- lmer(speed.per ~ authors.n + I(authors.n ^ 2) + (1|dataset), 
                   data = sampled_data,
                   weights = weight)
    
    ### extract weighted residuals
    m.s.quad.res <- 
      data.frame(res = residuals(m.s.quad),
                 weight = m.s.quad@frame$`(weights)`) %>%
      mutate(w.res = sqrt(weight*res^2))
  
    ## wilcox test: compare weighted residuals
    speed.quad <- 
      wilcox.test(m.s.lin.res$w.res, 
                  m.s.quad.res$w.res,
                  paired = TRUE)
  
    outcome_df$speed.quad <- speed.quad$p.value # record p-value
   
    ## BIC comparison
    outcome_df$speed.quad.BIC <- 
      BIC(m.s.lin) - BIC(m.s.quad)
    
    #######################
    # Step 4: Fit and evaluate citation models
    #######################
    ## linear modeling
    m.c.lin <- lmer(citation.per ~ authors.n + (1|dataset),
                    data = sampled_data,
                    weights = weight)
    
    ### extract weighted residuals
    m.c.lin.res <- data.frame(res = residuals(m.c.lin),
                              weight = m.c.lin@frame$`(weights)`) %>%
    mutate(w.res = sqrt(weight*res^2))
    
    ## log modeling
    m.c.log <- lmer(citation.per ~ log(authors.n) + (1|dataset), 
                    data = sampled_data,
                    weights = weight)
    
    ## extract weighted residuals
    m.c.log.res <- data.frame(res = residuals(m.c.log),
                              weight = m.c.log@frame$`(weights)`) %>%
      mutate(w.res = sqrt(weight*res^2))
    
    ## wilcox test: compare weighted residuals
    citation.log <- wilcox.test(m.c.lin.res$w.res, 
                                m.c.log.res$w.res,
                                paired = TRUE)
    
    outcome_df$citation.log <- citation.log$p.value # record p-value
    
    ## BIC comparison
    outcome_df$citation.log.BIC <- 
      BIC(m.c.lin) - BIC(m.c.log)
  
    #######################
    # Step 5: Fit and evaluate policy models
    #######################
    ## linear modeling
    m.p.lin <- lmer(policy.per ~ authors.n + (1|dataset),
                    data = sampled_data,
                    weights = weight)
  
    ## extract weighted residuals
    m.p.lin.res <- data.frame(res = residuals(m.p.lin),
                              weight = m.p.lin@frame$`(weights)`) %>%
      mutate(w.res = sqrt(weight*res^2))
    
    ## quadratic modeling
    m.p.quad <- lmer(policy.per ~ authors.n + I(authors.n ^ 2) + (1|dataset), 
                     data = sampled_data,
                     weights = weight)
    
    ## extract weighted residuals
    m.p.quad.res <- data.frame(res = residuals(m.p.quad),
                               weight = m.p.quad@frame$`(weights)`) %>%
      mutate(w.res = sqrt(weight*res^2))
  
    ## wilcox test: compare weighted residuals
    policy.quad <- wilcox.test(m.p.lin.res$w.res, 
                               m.p.quad.res$w.res,
                               paired = TRUE)
    
    outcome_df$policy.quad <- policy.quad$p.value # record p-value
  
    ## BIC comparison
    outcome_df$policy.quad.BIC <- 
      BIC(m.p.lin) - BIC(m.p.quad)
  
    #######################
    # Step 6: Fit and evaluate news models
    #######################
    ## linear modeling
    m.n.lin <- lmer(news.per ~ authors.n + (1|dataset),
                    data = sampled_data,
                    weights = weight)
  
    ## extract weighted residuals
    m.n.lin.res <- data.frame(res = residuals(m.n.lin),
                              weight = m.n.lin@frame$`(weights)`) %>%
      mutate(w.res = sqrt(weight*res^2))
    
    ## quadratic modeling
    m.n.quad <- lmer(news.per ~ authors.n + I(authors.n ^ 2) + (1|dataset), 
                     data = sampled_data,
                     weights = weight)
    
    ## extract weighted residuals
    m.n.quad.res <- data.frame(res = residuals(m.n.quad),
                               weight = m.n.quad@frame$`(weights)`) %>%
      mutate(w.res = sqrt(weight*res^2))
    
    ## wilcox test: compare weighted residuals
    news.quad <- wilcox.test(m.n.lin.res$w.res, 
                             m.n.quad.res$w.res,
                             paired = TRUE)
    
    outcome_df$news.quad <- news.quad$p.value # record p-value
  
    ## BIC comparison
    outcome_df$news.quad.BIC <- 
      BIC(m.n.lin) - BIC(m.n.quad)
  
  return(outcome_df)
}
```

# Power simulation
```{r}
# Determine the number of cores for parallelization
numCores <- 5 # set to 50 on final simulation
cl <- makeCluster(numCores)

# Export the dataframe and the power simulation function to each core
clusterExport(cl, varlist = c("DF", "sim"))

# Load the libraries used in the simulation in all cores
clusterEvalQ(cl, {
  library('tidyverse')
  library('lmerTest')
})

# choose number of iterations and sample sizes for simulation
iterations <- 10 # set to 1000 in final simulation
sample_sizes <- c(1000, 3300, 10000, 33000, 100000, 330000)

# create sample size list for lapply
sample_size_list = rep(sample_sizes, 
                       each = iterations)

# run  power simulation for each sample size in the list and binds all elements into a single dataframe
sim_outcome <- 
  bind_rows(parLapply(cl, 
                      sample_size_list, 
                      function(n)
                        sim(n = n, DF = DF)
                      )
            )

# free clusters after you are done simulating
stopCluster(cl)
```

# Optional: back up (or load) dataframe
```{r eval = FALSE}
# Optional: save dataframe after simulation
sim_outcome %>% 
  saveRDS(file.path('power_sim',
                    'btmobile25a_pwrsim_outcome.Rds')
          )

# Optional: load saved dataframe
sim_outcome <- 
  readRDS(file.path('power_sim',
                    'btmobile25a_pwrsim_outcome.Rds')
          )
```

# Plot power curve
Prepare dataframe
```{r}
# Check if the change in BIC is greater than 5
sim_outcome <- sim_outcome %>%
  mutate(
    speed.quad.test = 
      ifelse(speed.quad < 0.05, 
             1, 0),
    
    citation.log.test = 
      ifelse(citation.log < 0.05, 
             1, 0),
    
    policy.quad.test = 
      ifelse(policy.quad < 0.05, 
             1, 0),
    
    news.quad.test = 
      ifelse(news.quad < 0.05, 
             1, 0),
    
    speed.quad.BIC.test = 
      ifelse(speed.quad.BIC > 5, 
             1, 0),
    
    citation.log.BIC.test = 
      ifelse(citation.log.BIC > 5, 
             1, 0),
    
    policy.quad.BIC.test = 
      ifelse(policy.quad.BIC > 5, 
             1, 0),
    
    news.quad.BIC.test = 
      ifelse(news.quad.BIC > 5, 
             1, 0)
    )

# calculate the percentage of instances where decision criteria was met (and prep for viz)
power_df <- sim_outcome %>%
  
  # group by sample size
  group_by(sample_size) %>%
  
  # calculate power for Wilcox and BIC decision rules
  summarise(
    speed_quad = mean(speed.quad.test),
    citation_log = mean(citation.log.test),
    policy_quad = mean(policy.quad.test),
    news_quad = mean(news.quad.test),
    
    speed_quad_BIC = mean(speed.quad.BIC.test),
    citation_log_BIC = mean(citation.log.BIC.test),
    policy_quad_BIC = mean(policy.quad.BIC.test),
    news_quad_BIC = mean(news.quad.BIC.test)
    )  %>%
  
  # prepare for viz
  pivot_longer(cols = -sample_size,
               names_to = "model",
               values_to = "power") %>%
  
  mutate(
    outcome = case_when(
      model %in% c("speed_quad", "speed_quad_BIC") ~ "speed",
      model %in% c("citation_log", "citation_log_BIC") ~ "citation",
      model %in% c("policy_quad", "policy_quad_BIC") ~ "policy",
      model %in% c("news_quad", "news_quad_BIC") ~ "news"),
    
    fit = case_when(
      model %in% c("speed_quad", "speed_quad_BIC",
                   "policy_quad", "policy_quad_BIC", 
                   "news_quad", "news_quad_BIC") ~ "quad",
      model %in% c("citation_log", "citation_log_BIC") ~ "log"),
    
    comparison = case_when(
      model %in% c("speed_quad", "citation_log",
                   "policy_quad", "news_quad") ~ "wilcoxon",
      model %in% c("speed_quad_BIC", "citation_log_BIC",
                   "policy_quad_BIC", "news_quad_BIC") ~ "BIC")
    )
```

Plot power curve for Wilcoxon signed-rank test comparisons
```{r}
plot_w <- 
  
  # filter necessary outcomes
  power_df %>%
  filter(model %in% c("speed_quad", "citation_log", 
                      "policy_quad", "news_quad")) %>%
  
  # set aes
  ggplot(data = ., 
         aes(x = sample_size, 
             y = power, 
             color = model)) +
  
  # plot objects
  geom_line(size = 1, alpha = 0.7) +
  geom_point(size = 2, alpha = 0.7) +
  geom_hline(yintercept = 0.95, 
             linetype = "dotted", 
             color = "black") +
  
  # clean aes
  coord_cartesian(ylim = c(0, 330000)) +
  coord_cartesian(ylim = c(0, 1.04)) +
  theme_classic() +
  theme(legend.position = "none",
        axis.title.x = element_blank()) +
  labs(title = "Power Curve for Wilcoxon Test < 0.05", 
       color = "Model", 
       y = "Power") +
  scale_x_continuous(trans = "log10", 
                     labels = scales::label_number()) +
  scale_color_manual(
    values = c("speed_quad" = "#009E73", 
               "citation_log" = "#CC79A7",
               "policy_quad" = "#D55E00",
               "news_quad" = "#0072B2"))

```

Plot power curve for BIC comparisons
```{r}
plot_b <- 
  
  # filter necessary outcomes
  power_df %>%
  filter(model %in% c("speed_quad_BIC", 
                      "citation_log_BIC", 
                      "policy_quad_BIC", 
                      "news_quad_BIC")) %>%
  
  # set aes
  ggplot(data = ., 
         aes(x = sample_size, 
             y = power, 
             color = model)) +
  
  # plot objects
  geom_line(size = 1, alpha = 0.7) +
  geom_point(size = 2, alpha = 0.7) +
  geom_hline(yintercept = 0.95, 
             linetype = "dotted", 
             color = "black") +
  
  # fix aes
  coord_cartesian(ylim = c(0, 330000)) +
  coord_cartesian(ylim = c(0, 1.04)) +
  theme_classic() +
  theme(
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.title.x = element_blank()
  ) +
  labs(title = "Power Curve for \u0394 BIC < -5", y = "Power") +
  scale_x_continuous(trans = "log10", labels = scales::label_number()) +
  scale_color_manual(
    values = c("citation_log_BIC" = "#CC79A7", 
               "news_quad_BIC" = "#0072B2", 
               "policy_quad_BIC" = "#D55E00", 
               "speed_quad_BIC" = "#009E73"),
    labels = c("citation_log_BIC" = "DV = Scholarly article citations;
Model = Logarithmic", 
               "news_quad_BIC" = "DV = News citations;
Model = Quadratic", 
               "policy_quad_BIC" = "DV = Policy citations;
Model = Quadratic", 
               "speed_quad_BIC" = "DV = Response speed;
Model = Quadratic"),
    name = "Dependent Variable 
and Model Fit"
  )
```

Combine plots
```{r}
legend <- 
  get_legend(plot_b + theme(legend.position = "right"))

plot_b <- plot_b + theme(legend.position = "none")

plot_wb <- plot_grid(plot_w, plot_b, 
                     rel_widths = c(1, 1), ncol = 2)

plot_wb_label <- ggdraw() +
  draw_plot(plot_wb, 0, 0.05, 1, 0.95) +  # position and size of plot
  draw_text("Sample size (log scale)", x = 0.5, y = 0.02, fontface = "bold", size = 12)

plot_final <- plot_grid(plot_wb_label, legend, 
                        ncol = 2, rel_widths = c(2, 0.4))

print(plot_final)
```
